A Fast kNN Algorithm Using Multiple Space-Filling Curves

### 2022-04-12 2022/Sensors/sensors.tex
fix some typos
fix some errors

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; anton.shtanyuk@itmm.unn.ru (A.Sh.), sysoyev@vmk.unn.ru (A.S.)}
%$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
The paper considers a time-efficient implementation of the $k$ nearest neighbours (kNN) algorithm. A well-known approach for accelerating the kNN algorithm is to utilize dimensionality reduction methods based on the use of space-filling curves. In this paper, we take this approach further and propose an algorithm that employs multiple space-filling curves and is faster (with comparable quality) compared to the kNN algorithm which uses kd-trees to determine the nearest neighbors. A specific method for constructing multiple Peano curves is outlined, and statements are given about the preservation of object proximity information in the course of dimensionality reduction. An experimental comparison with known kNN implementations using kd-trees was performed using test and real-life data.
}

@@

The general scheme of object classification using kNN can be formulated as follows:
\begin{itemize}
\item calculate the distance from the object being classified to each of the objects in the training set;
\item select k objects of the training set with the minimum distance to them;
\item the class of the object being classified is the class most often found among the $k$ nearest neighbours.
\end{itemize}

@@

Let $\omega_i, 1 \leq i \leq S$, be the $i$-th object from the set of objects correlated with class $C_j$. Previously, a feature vector $y = (y_1, y_2, ..., y_N) \in R^N$, corresponding to each object under consideration was defined, i.e. 
\begin{equation}
\omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN})
\end{equation}

@@

\textbf{Remark}. The conditions of the theorem distinguish a specific neighborhood of the point $y^*$. This neighborhood comprises only such neighbours of this point which can be obtained by  shifting $y^*$ parallel to one of the coordinate axes by a distance not exceeding $2^{-p}$. By varying the value of $j, 1 \leq j \leq N$, under the conditions of the theorem, one can identify the nearest neighbours of point $y^*$ along $N$ coordinate directions. According to the statement, the proximity of points in the $N$-dimensional space in a particular direction will be reflected by the proximity of their preimages on one of the one-dimensional segments. In this case, the corresponding one-dimensional intervals can be different for different directions.

@@

The scheme of implementation of the kNN algorithm using multiple Peano curves will look as follows.
For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
\begin{equation}
\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \}
\end{equation} 
where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.

@@

Several datasets were synthesized to examine the search time for neighbours. First, DS-Random-1 and DS-Random-2 sets consisting of 100000 and 10000 objects, respectively, were generated.

@@

\begin{itemize}
\item \texttt{n\_samples} = 10000 (number of objects),
\item \texttt{n\_features} =  $N$ (dimensionality),
\item \texttt{n\_redundant} = 0 (number of redundant features),
\item \texttt{n\_informative} = $N - 1$  (number of informative features),
\item \texttt{n\_clusters\_per\_class} = 1 (number of clusters per class of objects generated).
\end{itemize}

@@

The number of objects of the first class is about 17000, the number of objects of the second class is 43000. The total number of objects examined is 60000.

>>>>

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
$^{1}$ \quad Department of Mathematical Software and Supercomputing Technologies, Lobachevsky University, 603950 Nizhny Novgorod, Russia; anton.shtanyuk@itmm.unn.ru (A.Sh.), alexander.sysoyev@itmm.unn.ru (A.S.)}
%$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
The paper considers a time-efficient implementation of the $k$ nearest neighbours (kNN) algorithm. A well-known approach for accelerating the kNN algorithm is to utilize dimensionality reduction methods based on the use of space-filling curves. In this paper, we take this approach further and propose an algorithm that employs multiple space-filling curves and is faster (with comparable quality) compared to the kNN algorithm which uses kd-trees to determine the nearest neighbours. A specific method for constructing multiple Peano curves is outlined, and statements are given about the preservation of object proximity information in the course of dimensionality reduction. An experimental comparison with known kNN implementations using kd-trees was performed using test and real-life data.
}

@@

The general scheme of object classification using kNN can be formulated as follows:
\begin{itemize}
\item calculate the distance from the object being classified to each of the objects in the training set;
\item select $k$ objects of the training set with the minimum distance to them;
\item the class of the object being classified is the class most often found among the $k$ nearest neighbours.
\end{itemize}

@@

Let $\omega_i, 1 \leq i \leq S$, be the $i$-th object from the set of objects correlated with class $C_j$. Previously, a feature vector $y = (y_1, y_2, ..., y_N) \in R^N$, corresponding to each object under consideration was defined, i.e. 
\begin{equation}
\omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN}).
\end{equation}

@@

\textbf{Remark}. The conditions of the theorem distinguish a specific neighbourhood of the point $y^*$. This neighbourhood comprises only such neighbours of this point which can be obtained by  shifting $y^*$ parallel to one of the coordinate axes by a distance not exceeding $2^{-p}$. By varying the value of $j, 1 \leq j \leq N$, under the conditions of the theorem, one can identify the nearest neighbours of point $y^*$ along $N$ coordinate directions. According to the statement, the proximity of points in the $N$-dimensional space in a particular direction will be reflected by the proximity of their preimages on one of the one-dimensional segments. In this case, the corresponding one-dimensional intervals can be different for different directions.

@@

The scheme of implementation of the kNN algorithm using multiple Peano curves will look as follows.
For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
\begin{equation}
\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \},
\end{equation} 
where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.

@@

Several datasets were synthesized to examine the search time for neighbours. First, DS-Random-1 and DS-Random-2 sets consisting of $100\;000$ and $10\;000$ objects, respectively, were generated.

@@

\begin{itemize}
\item \texttt{n\_samples} = $10\;000$ (number of objects),
\item \texttt{n\_features} =  $N$ (dimensionality),
\item \texttt{n\_redundant} = 0 (number of redundant features),
\item \texttt{n\_informative} = $N - 1$  (number of informative features),
\item \texttt{n\_clusters\_per\_class} = 1 (number of clusters per class of objects generated).
\end{itemize}

@@

The number of objects of the first class is about $17\;000$, the number of objects of the second class is $43\;000$. The total number of objects examined is $60\;000$.
